Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-10-01 review/response.Rmd

---
title: "Response to reviewer"
author: "Article ID: 2021-69"
output: pdf_document
---

@@

\textcolor{red}{The data set is first introduced in the subsection Simulation setup under Diagnosing an optimiser. To avoid confusion, we change the wording from "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}

-   "While the previous few examples have looked at the space of 1D bases ... 2D bases"

    -   I don't think the terms "1D bases" and "2D bases" are standard. Perhaps (orthonormal) "1-frames" and "2-frames"?

\textcolor{red}{We keep the term "1D bases" and "2D basis" in the paper since 1) "frame" is conventinally "basis", and 2) we use this term also in the software. To make this clearer, a sentence is added at the end of section "Projection pursuit guided tour" to note this.}

@@

\textcolor{red}{"details=TRUE" and "animate=TRUE" refers to the function "explore space pca". These two arguments are now mentioned in the implementation section under function "explore space pca", instead of subsection Diagnostic 3a and 3b.}

7.  In Figure 9, it seems CRS is doing better here than in Figure 8 -- maybe this could be mentioned.

\textcolor{red}{In the experiment of Figure 9, the parameter "max.tries" is set to 400 to allow guided tour to do its best instead of a default 25 as in Figure 8. A sentence in the Figure 9 captions is added to reflect this.}

8.  "not the case for all the indexes" missing a "." page 11 \textcolor{red}{Added.}

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.Rmd
last change to paper and response to reviewer

As a general visualisation method, tour produces animations of high-dimensional data via rotations of low-dimensional planes. There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview; manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection. Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection. Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path. Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames. Mathematical details of the geodesic interpolation can be found in @buja2005computational. The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].

@@

| $t$| $\mathbf{A}$| $I$| $S$| $j$| $l$| $V_{1}$| $V_{2}$| Notes|
|---|---|---|---|---|---|---|---|-------|
| $1$ | $\mathbf{A}_1$| $I_1$| $S_1$| 1| 1| $V_{11}$| $V_{12}$| start basis|
| $2$ | $\mathbf{A}_2$| $I_2$| $S_2$| 2| 1| $V_{21}$| $V_{22}$| search|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| 2     | $l_2$   | ...| ...| search (accepted) |
| ... | ... | ...| ...| 2     | 1       | ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| 2     | $k_2$   | ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| $J$   | 1       | ...| ...| search|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
| $T$ | $\mathbf{A}_T$ | $I_T$| $S_T$| $J$  | $l_J$| $V_{T1}$| $V_{T2}$| search (final)|
| ... | ... | ...| ...| $J$   | 1| ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
| ... | ... | ...| ...| $J$   | $k_J$| ...| ...| interpolation|
| ... | ... | ...| ...| $J+1$   | 1| ...| ...| search (last round)|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...    |
| $T^{\prime}$ | $\mathbf{A}_{T^{\prime}}$  | $I_{T^{\prime}}$ | $S_{T^{\prime}}$ | $J+1$ | $l_{J+1}$ | $V_{{T}^{\prime}1}$ | $V_{{T}^{\prime}2}$ | search (last round)   |

@@

While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames. Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space [@Buja1986-sw]. For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-torus}','\\@ref(fig:toy-torus-interactive)'))`. The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition. Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots. The final basis found by PD and CRS are shown in a larger shape and printed below respectively:

>>>>

As a general visualisation method, tour produces animations of high-dimensional data via rotations of low-dimensional planes. There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview; manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection. Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection. Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path. Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames. Mathematical details of the geodesic interpolation can be found in @buja2005computational. Notice that the term *frame* used in Buja's paper is conventionally called *basis*, which is used in this paper and the associated software. The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].

@@

| $t$| $\mathbf{A}$| $I$| $S$| $j$| $l$| $V_{1}$| $V_{2}$| Description|
|---|---|---|---|---|---|---|---|-------|
| $1$ | $\mathbf{A}_1$| $I_1$| $S_1$| 1| 1| $V_{11}$| $V_{12}$| start basis|
| $2$ | $\mathbf{A}_2$| $I_2$| $S_2$| 2| 1| $V_{21}$| $V_{22}$| search|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| 2     | $l_2$   | ...| ...| search (accepted) |
| ... | ... | ...| ...| 2     | 1       | ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| 2     | $k_2$   | ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
| ... | ... | ...| ...| $J$   | 1       | ...| ...| search|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
| $T$ | $\mathbf{A}_T$ | $I_T$| $S_T$| $J$  | $l_J$| $V_{T1}$| $V_{T2}$| search (final)|
| ... | ... | ...| ...| $J$   | 1| ...| ...| interpolation|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
| ... | ... | ...| ...| $J$   | $k_J$| ...| ...| interpolation|
| ... | ... | ...| ...| $J+1$   | 1| ...| ...| search (last round)|
| ... | ... | ...| ...| ...   | ...| ...| ...| ...    |
| $T^{\prime}$ | $\mathbf{A}_{T^{\prime}}$  | $I_{T^{\prime}}$ | $S_{T^{\prime}}$ | $J+1$ | $l_{J+1}$ | $V_{{T}^{\prime}1}$ | $V_{{T}^{\prime}2}$ | search (last round)   |

@@

While the previous few examples have looked at the space of 1D basis in a unit sphere, this section visualises the space of 2D basis. Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space [@Buja1986-sw]. For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-torus}','\\@ref(fig:toy-torus-interactive)'))`. The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition. Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots. The final basis found by PD and CRS are shown in a larger shape and printed below respectively:

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.tex
last change to paper and response to reviewer

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
\ref{fig:tour-path} shows a sketch of the tour path where the blue
frames are produced by the projection pursuit optimisation and the white
frames are interpolations between the blue frames. Mathematical details
of the geodesic interpolation can be found in
\citet{buja2005computational}. The aforementioned tour method has been
implemented in the R package \CRANpkg{tourr} \citep{tourr}.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.23}}@{}}
\toprule
\(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
\(V_{2}\) & Notes \\
\midrule
\endhead
\(1\) & \(\mathbf{A}_1\) & \(I_1\) & \(S_1\) & 1 & 1 & \(V_{11}\) &
\(V_{12}\) & start basis \\
\(2\) & \(\mathbf{A}_2\) & \(I_2\) & \(S_2\) & 2 & 1 & \(V_{21}\) &
\(V_{22}\) & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(l_2\) & \ldots{} &
\ldots{} & search (accepted) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & 1 & \ldots{} & \ldots{}
& interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(k_2\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T\) & \(\mathbf{A}_T\) & \(I_T\) & \(S_T\) & \(J\) & \(l_J\) &
\(V_{T1}\) & \(V_{T2}\) & search (final) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & \(k_J\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J+1\) & 1 & \ldots{} &
\ldots{} & search (last round) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T^{\prime}\) & \(\mathbf{A}_{T^{\prime}}\) & \(I_{T^{\prime}}\) &
\(S_{T^{\prime}}\) & \(J+1\) & \(l_{J+1}\) & \(V_{{T}^{\prime}1}\) &
\(V_{{T}^{\prime}2}\) & search (last round) \\
\bottomrule
\end{longtable}

@@

While the previous few examples have looked at the space of 1-frames in
a unit sphere, this section visualises the space of 2-frames. Recall
that the columns in a 2D basis are orthogonal to each other, so the
space of \(p \times 2\) bases is a torus in the \(p\)-D space
\citep{Buja1986-sw}. For \(p = 3\) one would see a classical 3D torus
shape as shown by the grey points in Figure \ref{fig:toy-torus}. The two
circles of the torus can be observed to be perpendicular to each other
and this can be linked back to the orthogonality condition. Two paths
from CRS and PD are plotted on top of the torus and coloured in green
and brown respectively to match the previous plots. The final basis
found by PD and CRS are shown in a larger shape and printed below
respectively:

>>>>

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
\ref{fig:tour-path} shows a sketch of the tour path where the blue
frames are produced by the projection pursuit optimisation and the white
frames are interpolations between the blue frames. Mathematical details
of the geodesic interpolation can be found in
\citet{buja2005computational}. Notice that the term \emph{frame} used in
Buja's paper is conventionally called \emph{basis}, which is used in
this paper and the associated software. The aforementioned tour method
has been implemented in the R package \CRANpkg{tourr} \citep{tourr}.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.23}}@{}}
\toprule
\(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
\(V_{2}\) & Description \\
\midrule
\endhead
\(1\) & \(\mathbf{A}_1\) & \(I_1\) & \(S_1\) & 1 & 1 & \(V_{11}\) &
\(V_{12}\) & start basis \\
\(2\) & \(\mathbf{A}_2\) & \(I_2\) & \(S_2\) & 2 & 1 & \(V_{21}\) &
\(V_{22}\) & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(l_2\) & \ldots{} &
\ldots{} & search (accepted) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & 1 & \ldots{} & \ldots{}
& interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(k_2\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T\) & \(\mathbf{A}_T\) & \(I_T\) & \(S_T\) & \(J\) & \(l_J\) &
\(V_{T1}\) & \(V_{T2}\) & search (final) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & \(k_J\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J+1\) & 1 & \ldots{} &
\ldots{} & search (last round) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T^{\prime}\) & \(\mathbf{A}_{T^{\prime}}\) & \(I_{T^{\prime}}\) &
\(S_{T^{\prime}}\) & \(J+1\) & \(l_{J+1}\) & \(V_{{T}^{\prime}1}\) &
\(V_{{T}^{\prime}2}\) & search (last round) \\
\bottomrule
\end{longtable}

@@

While the previous few examples have looked at the space of 1D basis in
a unit sphere, this section visualises the space of 2D basis. Recall
that the columns in a 2D basis are orthogonal to each other, so the
space of \(p \times 2\) bases is a torus in the \(p\)-D space
\citep{Buja1986-sw}. For \(p = 3\) one would see a classical 3D torus
shape as shown by the grey points in Figure \ref{fig:toy-torus}. The two
circles of the torus can be observed to be perpendicular to each other
and this can be linked back to the orthogonality condition. Two paths
from CRS and PD are plotted on top of the torus and coloured in green
and brown respectively to match the previous plots. The final basis
found by PD and CRS are shown in a larger shape and printed below
respectively:
